---
title: "Data Visualisation project"
author: "Janani Krishnamurthy"
date: "2024-12-13"
output: html_document
---

## Libraries

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(reshape2)
library(treemapify)
library(ggmosaic)
library(RColorBrewer)
```

## Data Pre-processing

```{r}
data <- read.csv("user_behavior_dataset 1.csv")
head(data)
```

```{r}
# Remove duplicates
data <- data %>% distinct()
summary(data)
nrow(data)
```

```{r}
# Standardize categorical data

# Convert `Operating System` and `Gender` columns to factor type
data$Operating.System <- as.factor(data$Operating.System)
data$Gender <- as.factor(data$Gender)

# Standardize case
data$Device.Model <- tolower(data$Device.Model)
data$Operating.System <- tolower(data$Operating.System)
data$Gender <- tolower(data$Gender)
```

```{r}
# Remove rows with unrealistic values
data <- data %>%
  filter(
    `App.Usage.Time..min.day.` >= 0 & `App.Usage.Time..min.day.` <= 1440, 
    `Screen.On.Time..hours.day.` >= 0 & `Screen.On.Time..hours.day.` <= 24,
    `Battery.Drain..mAh.day.` >= 0
  )
```

```{r}
# Remove rows with missing values
data <- data %>%
  drop_na()
```

```{r}
# Convert Age and User Behavior Class to integer
data$Age <- as.integer(data$Age)
data$User.Behavior.Class <- as.integer(data$User.Behavior.Class)
```

```{r}
#Renaming the column names

data <- data %>%
  rename(
    User_ID = User.ID,
    Device_Model = Device.Model,
    Operating_System = Operating.System,
    App_Usage_Time_Mins_Per_Day = App.Usage.Time..min.day.,
    Screen_Time_Hours_Per_Day = Screen.On.Time..hours.day.,
    Battery_Consumption_mAh_Per_Day = Battery.Drain..mAh.day.,
    Number_Of_Apps_Installed = Number.of.Apps.Installed,
    Data_Usage_MB_Per_Day = Data.Usage..MB.day.,
    User_Behavior_Class = User.Behavior.Class
  )
```

```{r}
summary(data)
```

```{r}
head(data)
```

```{r}
data <- data %>% mutate(Age_Group = cut(Age, breaks = seq(10, 70, by = 10), right = FALSE))
```


#### Heat Map

```{r}
library(reshape2)
num_vars <- data[, c("Screen_Time_Hours_Per_Day", 
                     "Battery_Consumption_mAh_Per_Day", 
                     "Data_Usage_MB_Per_Day", 
                     "App_Usage_Time_Mins_Per_Day", 
                     "Number_Of_Apps_Installed")]

# Create a simplified version of variable names
simplified_names <- c(
  "Screen Time (hrs/day)",
  "Battery Consumption (mAh/day)",
  "Data Usage (MB/day)",
  "App Usage Time (mins/day)",
  "Number of Apps Installed"
)

# Calculate correlation matrix
cor_matrix <- cor(num_vars, use = "complete.obs")

# Rename columns and rows in the correlation matrix
colnames(cor_matrix) <- simplified_names
rownames(cor_matrix) <- simplified_names

# Melt the correlation matrix for ggplot2
cor_melt <- melt(cor_matrix)

# Ensure both axes have the same factor levels
cor_melt$Var2 <- factor(cor_melt$Var2, levels = rev(colnames(cor_matrix)))

# Plot the heatmap
ggplot(cor_melt, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  geom_text(aes(label = round(value, 2)), color = "white", size = 4) +  
  scale_fill_gradient(low = "lightskyblue", high = "navy", name = "Correlation") +  
  labs(
    title = "Heatmap of Correlations between Key Variables",
    x = NULL,  
    y = NULL   
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),  
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12),  
    axis.text.y = element_text(size = 12),  
    panel.grid = element_blank(),  
    panel.border = element_blank(),  
    axis.line = element_line(color = "black")  
  )
```

#### Screen On Time

```{r}
# Calculate summary statistics for Screen On Time
box_summary <- data.frame(
  Metric = c("Mean", "Upper", "Lower"),
  Value = c(mean(data$Screen_Time_Hours_Per_Day, na.rm = TRUE),
            max(data$Screen_Time_Hours_Per_Day, na.rm = TRUE),
            min(data$Screen_Time_Hours_Per_Day, na.rm = TRUE)),
  y = c(14.5, 13.5, 12.5) 
)

# Add a constant x value since this is a single variable plot
box_summary$x <- 1

# Create the box plot with annotated text
ggplot(data, aes(y = Screen_Time_Hours_Per_Day)) +
  geom_boxplot(fill = "skyblue", color = "black", outlier.color = "red", outlier.shape = 16) + 
  geom_text(data = box_summary, aes(x = 0, y = y, label = paste0(Metric, ": ", round(Value, 2))),
            inherit.aes = FALSE, size = 4, color = "black") + 
  theme_minimal() +
  theme(
    panel.grid = element_blank(), 
    axis.line = element_line(color = "black") 
  ) +
  labs(
    title = "Box Plot of Screen On Time with Annotations",
    y = "Screen On Time (Hours per Day)"
  )
```


#### Gender

```{r}

gender_counts <- as.data.frame(table(data$Gender))
colnames(gender_counts) <- c("Gender", "Count")

# Calculate percentages
gender_counts$Percentage <- round((gender_counts$Count / sum(gender_counts$Count)) * 100, 1)

# Create the polar graph with percentage labels
ggplot(gender_counts, aes(x = "", y = Count, fill = Gender)) +
  geom_bar(stat = "identity", width = 1, color = "black") + 
  coord_polar(theta = "y") + 
  scale_fill_manual(values = c("male" = "skyblue", "female" = "pink")) + 
  geom_text(aes(label = paste0(Percentage, "%")), 
            position = position_stack(vjust = 0.5), color ="black", size = 5) + 
  labs(
    title = "Gender Distribution",
    fill = "Gender"
  ) +
  theme_void() + 
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5) 
  )
```

#### User Behaviour Class

```{r}
user_class_distribution <- data %>%
  count(User_Behavior_Class) %>%
  mutate(Percentage = round(n / sum(n) * 100, 1))  

# Create a pie chart
ggplot(user_class_distribution, aes(x = "", y = Percentage, fill = as.factor(User_Behavior_Class))) +
  geom_bar(stat = "identity", width = 1, color = "white") +  
  coord_polar(theta = "y") +  
  labs(
    title = "User Behavior Class Distribution",
    fill = "Behavior Class"
  ) +
  geom_text(aes(label = paste0(Percentage, "%")), 
            position = position_stack(vjust = 0.5), size = 4, color = "black") +  
  scale_fill_manual(values = c("lightblue", "skyblue", "dodgerblue", "deepskyblue", "steelblue")) + 
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.title = element_blank(),  
    axis.text = element_blank(),  
    panel.grid = element_blank(),  
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )
```


#### User Behaviour Class by App Usage and Gender (Volin Plot)

```{r}
# Violin Plot for App Usage Time Distribution by Gender and User Class
ggplot(data, aes(x = as.factor(User_Behavior_Class), y = App_Usage_Time_Mins_Per_Day, fill = Gender)) +
  geom_violin(alpha = 0.8, color = "black") + 
  labs(
    title = "App Usage Time Distribution by Gender and User Class",
    x = "User Behavior Class",
    y = "App Usage Time (Minutes/Day)",
    fill = "Gender"
  ) +
  scale_fill_manual(values = c("male" = "skyblue", "female" = "pink")) + 
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.position = "top", 
    panel.grid = element_blank(), 
    axis.line = element_line(color = "black") 
  )
```


#### Same using Box plot for seeing actual difference between Gender

```{r}
# Box Plot
ggplot(data, aes(x = as.factor(User_Behavior_Class), y = App_Usage_Time_Mins_Per_Day, fill = Gender)) +
  geom_boxplot(alpha = 0.6, position = position_dodge(0.9), outlier.shape = NA) +
  labs(
    title = "App Usage Time Distribution by Gender and User Class",
    x = "User Behavior Class",
    y = "App Usage Time (Minutes/Day)",
    fill = "Gender",
    color = "Gender"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.position = "top"
  ) +
  scale_fill_manual(values = c("pink", "lightblue"))
```


#### Bar Graph for the Average App Usage Time by Gender and User Behavior Class 

```{r}
bar_data <- data %>%
  group_by(User_Behavior_Class, Gender) %>%
  summarise(Average_App_Usage = mean(App_Usage_Time_Mins_Per_Day, na.rm = TRUE)) %>%
  ungroup()

ggplot(bar_data, aes(x = as.factor(User_Behavior_Class), y = Average_App_Usage, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge", color = "black", alpha = 0.9, width = 0.7) + 
  geom_text(aes(label = round(Average_App_Usage, 1)), 
            position = position_dodge(0.7), vjust = 1.5, hjust= 0.5, size = 4, color = "black") +  
  labs(
    title = "Average App Usage Time by Gender and User Behavior Class",
    x = "User Behavior Class",
    y = "Average App Usage Time (Minutes/Day)",
    fill = "Gender"
  ) +
  scale_fill_manual(values = c("male" = "skyblue", "female" = "pink")) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 600)) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.position = "top", 
    panel.grid = element_blank(), 
    axis.line = element_line(color = "black") 
  )
```

#### Average Data Usage by User Behaviour Class

```{r}
avg_data_usage <- data %>%
  group_by(User_Behavior_Class) %>%
  summarise(Average_Data_Usage = round(mean(Data_Usage_MB_Per_Day, na.rm = TRUE), 1))

# Line plot
ggplot(avg_data_usage, aes(x = as.factor(User_Behavior_Class), y = Average_Data_Usage, group = 1)) +
  geom_line(color = "skyblue", size = 1.2) +
  geom_point(color = "blue", size = 3) +
  geom_text(aes(label = Average_Data_Usage), vjust = 0.4, hjust = -0.3, size = 4) +
  labs(
    title = "Average Data Usage by User Behavior Class",
    x = "User Behavior Class",
    y = "Average Data Usage (MB/Day)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    panel.grid = element_blank(), 
    axis.line = element_line(color = "black")
  )
```

#### Average Number of Apps Installed by User Behaviour Class and Gender

```{r}
# Summarize the data for the average number of apps installed by User Behavior Class and Gender
avg_apps_installed_gender <- data %>%
  group_by(User_Behavior_Class, Gender) %>%
  summarise(Average_Apps_Installed = round(mean(Number_Of_Apps_Installed, na.rm = TRUE), 0)) %>%
  mutate(User_Behavior_Class = as.factor(User_Behavior_Class))

# Plot the average number of apps installed by User Behavior Class and Gender
ggplot(avg_apps_installed_gender, aes(x = User_Behavior_Class, y = Average_Apps_Installed, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge", color = "black", alpha = 0.9, width = 0.7) +
  geom_text(aes(label = Average_Apps_Installed), position = position_dodge(width = 0.7), vjust = -0.5, size = 4, color = "black") +
  scale_fill_manual(values = c("female" = "pink", "male" = "skyblue")) +
  labs(
    title = "Average Number of Apps Installed by User Behavior Class and Gender",
    x = "User Behavior Class",
    y = "Average Number of Apps Installed",
    fill = "Gender"
  ) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 100)) + 
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    panel.grid = element_blank(), 
    axis.line = element_line(color = "black"), 
    legend.position = "top"
  )
```

#### Screen Time by User Behaviour Class

```{r}
ggplot(data, aes(x = Screen_Time_Hours_Per_Day, fill = as.factor(User_Behavior_Class))) +
  geom_density(alpha = 0.6) +  
  scale_fill_brewer(palette = "Set3") +  
  labs(
    title = "Density of Screen Time by User Behavior Class",
    x = "Screen Time (Hours/Day)",
    y = "Density",
    fill = "Behavior Class"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.position = "right",
    panel.grid = element_blank(),  
    axis.line = element_line(color = "black")  
  )
```

#### Average Data Usage by User Behaviour Class and Age Group

```{r}
heatmap_data <- data %>%
  group_by(User_Behavior_Class, Age_Group) %>%
  summarise(Average_Data_Usage = round(mean(Data_Usage_MB_Per_Day, na.rm = TRUE), 1))

# Heatmap
ggplot(heatmap_data, aes(x = Age_Group, y = as.factor(User_Behavior_Class), fill = Average_Data_Usage)) +
  geom_tile(color = "white") +
  geom_text(aes(label = Average_Data_Usage), size = 4, color = "black") +
  scale_fill_gradient(low = "lightblue", high = "blue", name = "Data Usage (MB)") +
  labs(
    title = "Heatmap of Average Data Usage by User Behavior Class and Age Group",
    x = "Age Group",
    y = "User Behavior Class"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    panel.grid = element_blank() 
  )
```

#### User Behaviour Class relationship with all Variaables

```{r}
# Select relevant variables
scatter_data <- data[, c("User_Behavior_Class", 
                         "Screen_Time_Hours_Per_Day", 
                         "Battery_Consumption_mAh_Per_Day", 
                         "Data_Usage_MB_Per_Day", 
                         "App_Usage_Time_Mins_Per_Day", 
                         "Number_Of_Apps_Installed")]

# Convert data to long format for faceting
scatter_long <- pivot_longer(scatter_data, 
                             cols = -User_Behavior_Class, 
                             names_to = "Variable", 
                             values_to = "Value")

# Plot scatterplots with faceting
ggplot(scatter_long, aes(x = Value, y = User_Behavior_Class, color = Variable)) +
  geom_point(alpha = 0.6) +
  facet_wrap(~ Variable, scales = "free_x") +
  labs(title = "Scatterplots of User Behavior Class vs Other Variables",
       x = "Variable Value",
       y = "User Behavior Class") +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

#### Battery Usage vs Data Usage by User Behaviour Class and Gender

```{r}
ggplot(data, aes(x = Data_Usage_MB_Per_Day, y = Battery_Consumption_mAh_Per_Day, color = Gender)) +
  geom_point(alpha = 0.6) +
  facet_wrap(~ User_Behavior_Class,scales = "free_x") +
  labs(title = "Battery Consumption vs. Data Usage by User Class and Gender",
       x = "Data Usage (MB/day)",
       y = "Battery Consumption (mAh/day)",
       color = "Gender") +
  theme_minimal() +
    theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5), 
    strip.text = element_text(size = 12, face = "bold"), 
    axis.line = element_line(color = "black"), 
    panel.grid = element_blank(), 
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 10), 
    axis.text.y = element_text(size = 10), 
    legend.position = "right" 
  )
```


#### Initial Graphs :Created by Me. But is not used as such in the report or PPT

```{r}
# Bar Plot of Device Models
ggplot(data, aes(x = reorder(Device_Model, Device_Model, function(x) -length(x)))) +
  geom_bar(fill = "orange", alpha = 0.7) +
  geom_text(stat = "count", aes(label = ..count..), vjust = 1.5, color = "black") +
  labs(title = "Device Model Distribution", x = "Device Model", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
ggplot(data, aes(x = reorder(Device_Model, Device_Model, function(x) -length(x)), fill = Gender)) +
  geom_bar(position = "dodge", alpha = 0.7) +
  labs(title = "Device Model Distribution by Gender", x = "Device Model", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r}
# Load necessary libraries
library(dplyr)
library(ggplot2)

# Calculate total count per device model
total_counts <- data %>%
  group_by(Device_Model) %>%
  summarise(Total = n())

# Create the grouped bar chart with total count labels
ggplot(data, aes(x = reorder(Device_Model, -table(Device_Model)[Device_Model]), fill = Gender)) +
  geom_bar(position = "dodge", alpha = 0.7) +
  geom_text(data = total_counts, aes(x = Device_Model, y = Total + 5, label = Total), 
            inherit.aes = FALSE, color = "black", size = 3, vjust = -0.5) + 
  labs(title = "Device Model Distribution by Gender with Total Count", 
       x = "Device Model", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


```{r}
ggplot(data, aes(x = reorder(Device_Model, -table(Device_Model)[Device_Model]), fill = Gender)) +
  geom_bar(position = "dodge", alpha = 0.7) +
  geom_text(stat = "count", aes(label = ..count..), 
            position = position_dodge(width = 0.9), vjust = -0.5, color = "black", size = 3) +
  labs(title = "Device Model Distribution by Gender", 
       x = "Device Model", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r}
# Scatter plot for Number of Apps Installed vs. Battery Drain
ggplot(data, aes(x = Number_Of_Apps_Installed, y = Battery_Consumption_mAh_Per_Day)) +
  geom_point(color = "blue") +
  labs(
    title = "Number of Apps Installed vs. Battery Drain",
    x = "Number of Apps Installed",
    y = "Battery Drain (mAh/day)"
  ) +
  theme_minimal()
```


# Homework 7 

#Visulaising User Behavior class using bar and pie chart

```{r}
# Bar Plot for User_Behavior_Class
ggplot(data, aes(x = User_Behavior_Class)) +
  geom_bar(fill = "skyblue", color = "black") +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5) +
  labs(
    title = "Distribution of User Behavior Class",
    x = "User Behavior Class",
    y = "Count"
  ) +
  theme_minimal()

```


```{r}
# Pie Chart for User_Behavior_Class
behavior_counts <- data %>%
  dplyr::count(User_Behavior_Class) %>%
  dplyr::mutate(percentage = n / sum(n) * 100)

ggplot(behavior_counts, aes(x = "", y = percentage, fill = as.factor(User_Behavior_Class))) +
  geom_col(width = 1, color = "white") +
  coord_polar(theta = "y") +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), 
            position = position_stack(vjust = 0.5), color = "white") +
  scale_fill_brewer(palette = "Set1") +  
  labs(
    title = "User Behavior Class Distribution (Pie Chart)",
    x = NULL,
    y = NULL,
    fill = "User Behavior Class"
  ) +
  theme_void() +
  theme(legend.position = "right")

```

#Visualising Screen_Time_Hours_Per_Day

```{r}
ggplot(data, aes(x = Screen_Time_Hours_Per_Day)) +
  geom_histogram(binwidth = 0.5, fill = "skyblue", color = "black") +
  labs(
    title = "Distribution of Screen Time per Day",
    x = "Screen Time (Hours per Day)",
    y = "Frequency"
  ) +
  theme_minimal()

```


```{r}
ggplot(data, aes(x = Screen_Time_Hours_Per_Day)) +
  geom_density(fill = "blue", alpha = 0.5) +
  labs(
    title = "Density of Screen Time (Hours per Day)",
    x = "Screen Time (Hours per Day)",
    y = "Density"
  ) +
  theme_minimal()
```

#Visualising Gender


```{r}
ggplot(data, aes(x = Gender)) +
  geom_bar(fill = "skyblue") +
  labs(title = "Gender Distribution of Mobile Users", x = "Gender", y = "Count") +
  theme_minimal()
```


#Visualising Age

```{r}
 #2. Age Distribution - Histogram
ggplot(data, aes(x = Age)) +
  geom_histogram(binwidth = 5, fill = "white", color = "black") +
  labs(title = "Age Distribution of Mobile Users", x = "Age", y = "Frequency") +
  theme_minimal()
```

#Visualising Battery Drain

```{r}
# 3. Battery Drain (mAh/day) - Density Plot
ggplot(data, aes(x = `Battery_Consumption_mAh_Per_Day`)) +
  geom_density(fill = "salmon", alpha = 0.7) +
  labs(title = "Battery Drain Distribution", x = "Battery Drain (mAh/day)", y = "Density") +
  theme_minimal()
```

#Visualising Device Model

```{r}
# Get unique values of Device.Model
unique_models <- unique(data$Device_Model)
print(unique_models)

```

```{r}
# Bar chart for Device Model

ggplot(data, aes(x = Device_Model)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Distribution of Device Models", x = "Device Model", y = "Count")

```

#Visualising Operating System

```{r}
# Pie chart for Operating System
os_counts <- table(data$Operating_System)
pie(os_counts, main = "Distribution of Operating Systems", col = rainbow(length(os_counts)))

```

#Visualising App_Usage_Time_Mins_Per_Day

```{r}
# Histogram for App Usage Time
ggplot(data, aes(x = App_Usage_Time_Mins_Per_Day)) +
  geom_histogram(binwidth = 30, color = "black", fill = "blue", alpha = 0.7) +
  labs(title = "Distribution of App Usage Time per Day", x = "App Usage Time (Minutes)", y = "Frequency")

```

#Combination of Screen Time and User Class

```{r}
ggplot(data, aes(x = Screen_Time_Hours_Per_Day, fill = as.factor(User_Behavior_Class))) +
  geom_density(alpha = 0.6) +
  labs(
    title = "Density of Screen Time per Day by User Behavior Class",
    x = "Screen Time (Hours per Day)",
    y = "Density",
    fill = "User Behavior Class"
  ) +
  theme_minimal()

```


```{r}
# Jitter Plot of Screen Time by User Behavior Class
ggplot(data, aes(x = as.factor(User_Behavior_Class), y = Screen_Time_Hours_Per_Day, color = as.factor(User_Behavior_Class))) +
  geom_jitter(width = 0.1, alpha = 0.5) +
  labs(
    title = "Individual Screen Time Data Points by User Behavior Class",
    x = "User Behavior Class",
    y = "Screen Time (Hours per Day)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```

#Homework 9:

#Graph 1: 
This graph helps analyze if battery consumption correlates with data usage and whether the pattern differs by gender within each user behavior class. High battery consumption with low data usage may indicate inefficient apps or background activity. <br/>

```{r}
ggplot(data, aes(x = Data_Usage_MB_Per_Day, y = Battery_Consumption_mAh_Per_Day, color = Gender)) +
  geom_point(alpha = 0.6) +
  facet_wrap(~ User_Behavior_Class,scales = "free_x") +
  labs(title = "Battery Consumption vs. Data Usage by User Class and Gender",
       x = "Data Usage (MB/day)",
       y = "Battery Consumption (mAh/day)",
       color = "Gender") +
  theme_minimal()
```


#Graph 2:

This graph shows the proportion of each device model within different user behavior classes. It helps identify if specific user behavior classes prefer certain devices more than others. <br/>

```{r}
ggplot(data, aes(x = as.factor(User_Behavior_Class), fill = Device_Model)) +
  geom_bar(position = "fill") +  # 'fill' shows proportions
  labs(title = "Proportion of Device Models by User Behavior Class",
       x = "User Behavior Class",
       y = "Proportion",
       fill = "Device Model") +
  theme_minimal()
```



#Graph 3 :

The violin plot will show the distribution and density of app usage time for each gender, separated by user behavior class. It helps identify if certain classes or genders spend more time using apps and how usage varies within each class.. <br/>

```{r}
ggplot(data, aes(x = Gender, y = App_Usage_Time_Mins_Per_Day, fill = as.factor(User_Behavior_Class))) +
  geom_violin(alpha = 0.6) +
  labs(title = "App Usage Time Distribution by Gender and User Class",
       x = "Gender",
       y = "App Usage Time (Minutes/Day)",
       fill = "User Class") +
  theme_minimal()

```

#Graph 4 :

```{r}
library(ggplot2)
library(reshape2)

# Select numerical variables
num_vars <- data[, c("Screen_Time_Hours_Per_Day", "Battery_Consumption_mAh_Per_Day", 
                     "Data_Usage_MB_Per_Day", "App_Usage_Time_Mins_Per_Day", 
                     "Number_Of_Apps_Installed")]

# Calculate correlation matrix
cor_matrix <- cor(num_vars, use = "complete.obs")

# Melt for ggplot2
cor_melt <- melt(cor_matrix)

cor_melt$Var2 <- factor(cor_melt$Var2, levels = rev(colnames(cor_matrix)))


# Plot heatmap
ggplot(cor_melt, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  labs(title = "Heatmap of Correlations between Key Variables", fill = "Correlation") +
  theme_minimal() +
  scale_fill_gradient(low = "lightskyblue", high = "navy") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  
    axis.text.y = element_text(angle = 0)  
  )

```



#Graph 5 :

This visual helps identify whether certain user behavior classes favor iOS or Android more. For example, if a class heavily uses iOS, it may reflect a preference for specific usage patterns or demographics. <br/>


```{r}
unique(data$Operating_System)
```

```{r}
library(ggplot2)
library(dplyr)

# Clean the data to ensure no NA or empty values
data_clean <- data %>%
  filter(!is.na(Operating_System) & Operating_System != "")

# Summarize data for OS distribution within each user behavior class
os_summary <- data_clean %>%
  group_by(User_Behavior_Class, Operating_System) %>%
  summarise(Count = n(), .groups = "drop")

# Plot bar graph with counts and adjusted label position
ggplot(os_summary, aes(x = Operating_System, y = Count, fill = Operating_System)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.6) +  
  geom_text(aes(label = Count), 
            vjust = 1.5,  
            size = 4, 
            color = "white") +  
  facet_wrap(~ User_Behavior_Class, scales = "free_y") + 
  labs(title = "Operating System Distribution by User Behavior Class", 
       y = "Count", x = "Operating System", fill = "Operating System") +
  theme_minimal() +
  theme(legend.position = "none",  
        strip.text = element_text(size = 10),  
        axis.text.x = element_text(angle = 45, hjust = 1))  


```

```{r}
library(dplyr)

# Clean the data to ensure no NA or empty values
data_clean <- data %>%
  filter(!is.na(Operating_System) & Operating_System != "")

# Summarize data for OS distribution within each user behavior class
os_summary <- data_clean %>%
  group_by(User_Behavior_Class, Operating_System) %>%
  summarise(Count = n(), .groups = "drop")

# Calculate total count per User Behavior Class to compute percentages
os_percentage <- os_summary %>%
  group_by(User_Behavior_Class) %>%
  mutate(Total_Count = sum(Count)) %>%
  ungroup() %>%
  mutate(Percentage = (Count / Total_Count) * 100)

# Display the table with percentages
os_percentage_table <- os_percentage %>%
  select(User_Behavior_Class, Operating_System, Count, Percentage) %>%
  arrange(User_Behavior_Class, Operating_System)

# Print the table
print(os_percentage_table)

```

```{r}
library(ggplot2)

# Create a pie chart with better label placement
ggplot(os_percentage_table, aes(x = "", y = Percentage, fill = Operating_System)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  facet_wrap(~ User_Behavior_Class) + 
  labs(title = "Operating System Distribution by User Behavior Class", fill = "Operating System") +
  theme_void() +  
  theme(
    plot.title = element_text(hjust = 0.5), 
    strip.text = element_text(size = 12)  
  ) +
  # Adding percentage labels
  geom_text(aes(label = paste0(round(Percentage, 1), "%")), 
            position = position_stack(vjust = 0.6),  
            size = 4, 
            color = "white")   


```






```{r}
table(data_clean$User_Behavior_Class, data_clean$Operating_System)
```

#Graph 6 :

```{r}
ggplot(data, aes(x = factor(User_Behavior_Class), y = Data_Usage_MB_Per_Day, fill = factor(User_Behavior_Class))) +
  geom_bar(stat = "identity") +
  labs(title = "Data Usage by User Behavior Class",
       x = "User Behavior Class",
       y = "Data Usage (MB/Day)",
       fill = "User Class") +
  theme_minimal()
```


```{r}
library(ggplot2)
library(tidyr)

# Select relevant variables
scatter_data <- data[, c("User_Behavior_Class", 
                         "Screen_Time_Hours_Per_Day", 
                         "Battery_Consumption_mAh_Per_Day", 
                         "Data_Usage_MB_Per_Day", 
                         "App_Usage_Time_Mins_Per_Day", 
                         "Number_Of_Apps_Installed")]

# Convert data to long format for faceting
scatter_long <- pivot_longer(scatter_data, 
                             cols = -User_Behavior_Class, 
                             names_to = "Variable", 
                             values_to = "Value")

# Plot scatterplots with faceting
ggplot(scatter_long, aes(x = Value, y = User_Behavior_Class, color = Variable)) +
  geom_point(alpha = 0.6) +
  facet_wrap(~ Variable, scales = "free_x") +
  labs(title = "Scatterplots of User Behavior Class vs Other Variables",
       x = "Variable Value",
       y = "User Behavior Class") +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))



```






